---
title: "Untitled"
author: "Zachary M. Smith"
date: "July 25, 2018"
output: html_document
---

## Parametric Tests: Reach by Year

Generate a metric key. This data frame will represent sample information, like `reach`, `bank`, and `year`, that the metric values will be appended to.
```{r}
metrics.key <- standard.df %>% 
  select(unique_id, reach_year, reach, bank, year) %>% 
  distinct()
```

### Calculate Metrics

Calculate the metrics of interest using the __mmir__ package functions.
```{r}
metrics.df <- metrics.key %>% 
  mutate(rich = taxa_rich(standard.df, unique_id, genus),
         rich_ephemeroptera = taxa_rich(standard.df, unique_id, order, genus, "ephemeroptera"),
         rich_trichoptera = taxa_rich(standard.df, unique_id, order, genus, "trichoptera"),
         rich_diptera = taxa_rich(standard.df, unique_id, order, genus, "diptera"),
         pct_chironomidae = taxa_pct(standard.df, unique_id, count,
                                     taxon.col = family, taxon = "chironomidae"),
         pct_corbiculidae = taxa_pct(standard.df, unique_id, count,
                                     taxon.col = family, taxon = "corbiculidae"),
         pct_elmidae = taxa_pct(standard.df, unique_id, count,
                                taxon.col = family, taxon = "elmidae"),
         pct_hydropsychidae = taxa_pct(standard.df, unique_id, count,
                                       taxon.col = family, taxon = "hydropsychidae"),
         pct_ephemeroptera = taxa_pct(standard.df, unique_id, count,
                                      taxon.col = order, taxon = "ephemeroptera"),
         pct_plecoptera = taxa_pct(standard.df, unique_id, count,
                                   taxon.col = order, taxon = "plecoptera"),
         pct_trichoptera = taxa_pct(standard.df, unique_id, count,
                                    taxon.col = order, taxon = "trichoptera"),
         pct_ept = taxa_pct(standard.df, unique_id, count,
                            taxon.col = order, taxon = c("ephemeroptera", 
                                                         "plecoptera",
                                                         "trichoptera"))
  )
```

Identify the metrics of interest and the reaches of interest.
```{r}
metrics.vec <- c("rich", "rich_ephemeroptera", "rich_trichoptera",
                 "rich_diptera", "pct_chironomidae",
                 "pct_corbiculidae",
                 "pct_elmidae", "pct_hydropsychidae", "pct_ephemeroptera",
                 "pct_plecoptera", "pct_trichoptera", "pct_ept")
reach.vec <- unique(metrics.df$reach)
```

### Kruskal-Wallis and post hoc Dunn's Test

For each metric, run the Kruskal-Wallis and post hoc Dunn's test to compare the values from each reaches collection year. The values are stored in a dataframe. Keep in mind that the Kruskal-Wallis p-value (`kruskal_p`) must be < α before it is appropriate to review the Dunn's test adjusted p-value (`dunn_adjusted_p`).
```{r}
dunn.df <- map_df(metrics.vec, function(metric.i){
  map_df(reach.vec, function(reach.i) {
    metrics.sub <- metrics.df %>% 
      filter(reach  == reach.i) %>% 
      mutate(reach_year = as.factor(reach_year)) %>% 
      rename(metric = metric.i)
    
    kruskal.df <- kruskal.test(metric ~ as.factor(reach_year), data = metrics.sub) %>%
      broom::tidy() %>%
      mutate(metric = metric.i,
             reach = reach.i) %>%
      select(reach, metric, everything())
    
    final.df <- dunn.test::dunn.test(metrics.sub$metric, metrics.sub$reach_year)
    final.df <- tibble(
      metric = metric.i,
      kruskal_p = kruskal.df$p.value,
      kruskal_stat = kruskal.df$statistic,
      dunn_comparisons = final.df$comparisons,
      dunn_p_adjusted = final.df$P.adjusted,
      dunn_p_value = final.df$P,
      dunn_z_score = final.df$Z,
      dunn_chi2 = final.df$chi2
    )
  })
})

DT::datatable(dunn.df)

```

### Boxplots

Transform the dataframe from a wide data format to a long data format.
```{r}
long.df <- metrics.df %>% 
  gather(metric, value, rich:pct_ept)
```

Generate boxplots for each reach and metric. The values are grouped by collection year.
```{r, fig.width=8, fig.height=20}
long.df %>% 
  ggplot(aes(as.factor(year), value)) +
  geom_boxplot() +
  facet_wrap(~ reach + metric, ncol = 3, scales = "free")
```

### Percentiles of Interest

Calculate the percentiles of interest for each reach and metric. Additionally, a `lower_fence` and `upper_fence` are calculated by multiplying the interquartile range by 1.5 and then either subtracting this values from the 25th percentile or adding this value to the 75th percentile.
```{r}
percentiles.df <- long.df %>% 
  group_by(reach, metric) %>% 
  summarize(iqr = IQR(value),
            quant10 = quantile(value, 0.10),
            quant25 = quantile(value, 0.25),
            quant75 = quantile(value, 0.75),
            quant90 = quantile(value, 0.90),
            lower_fence = quant25 - iqr * 1.5,
            upper_fence = quant75 + iqr * 1.5
  ) %>% 
  ungroup() %>% 
  select(reach, metric, lower_fence, quant10, quant25, quant75, quant90, upper_fence)

DT::datatable(percentiles.df)
```

